---
title: "Data Limited Methods of Stock Assessment"
author: "Dustin Duncan"
format: 
  html:
    code-fold: true
    toc: true
    number-sections: true
    embed-resources: true
editor: visual
execute:
  echo: true
  message: false
  warning: false
---


## Data limited (1st iteration)

Anderson, S. C., Afflerbach, J., Cooper, A. B., Dickey-Collas, M., Jensen, O. P.,
Kleisner, K. M., et al. (2016) Datalimited: Stock Assessment Methods for Data-
Limited Fisheries. R Package Version 0.1.0. Available at: https://github.com/
datalimited/datalimited.

[Github link](https://github.com/datalimited/datalimited)

```{r}
library("datalimited")
set.seed(1)
?cmsy
bg =blue_gren

x <- cmsy(blue_gren$yr, ct = blue_gren$ct, reps = 2e4)
names(x)
#> [1] "theta"       "biomass"     "bmsy"        "msy"         "mean_ln_msy"
#> [6] "bbmsy"
par(mfrow = c(2, 2))
plot(blue_gren$yr, blue_gren$ct, type = "o", xlab = "Year", 
  ylab = "Catch (t)")
plot(blue_gren$yr,  apply(x$biomass, 2, median)[-1], type = "o",
  ylab = "Estimated biomass", xlab = "Year")
hist(x$bmsy)
plot(x$theta$r, x$theta$k, col = "#00000030")

?comsir
x <- comsir(ct = blue_gren$ct, yr = blue_gren$yr, nsim = 1e5,
  n_posterior = 2e3)
par(mfrow = c(1, 2))
hist(x$quantities$bbmsy)
with(x$posterior, plot(r, k))
```


## Data limited 2 (catch-only methods of assessment)

Free, C. M. (2018) Datalimited2: More Stock Assessment Methods for Data-Limited
Fisheries. R Package Version 0.1.0. Available at: https://github.com/cfree14/
datalimited2.

[Github Link](https://github.com/cfree14/datalimited2)

The package implements the following catch-only stock assessment models:

* cMSY from Froese et al. 2017; see ?cmsy2

* Bayesian surplus production model (BSM) from Froese et al. 2017; see ?bsm

* Boosted regression tree approach (zBRT) from Zhou et al. 2017a; see ?zbrt

* Optimized catch-only model (OCOM) from Zhou et al. 2017b; see ?ocom

* Refined ORCS approach (rORCS) from Free et al. 2017; see ?rorcs

* Multispecies cMSY (MS-cMSY) from Free et al. in prep; see ?ms_cmsy

* The results of all but the rORCS model can be visualized using the ?plot_dlm function.

* The performance of catch-only methods (i.e., their ability to accurately estimate B/BMSY or accurately classify categorical status) can be compared using the ?performance function.

* Many of these catch-only methods require estimates of resilience to set priors and resilience estimates can be retrieved from FishBase/SeaLifeBase busing the ?resilience function.

```{r}
# devtools::install_github("cfree14/datalimited2")
library(datalimited2)

?resilience

```

## Random Forest method 

Liaw, A. and Wiener, M. (2002). Classification and Regression by Randomforest.
R News. 2, 18–22.

```{r}
# install.packages("randomForest")
library(randomForest)

```

## FishLife (critical information about fish taxa)

Thorson, J. T. (2020). Predicting recruitment density dependence and intrinsic growth rate for all fishes worldwide using a data‐integrated life‐history model. Fish and Fisheries, 21(2), 237-251.

```{r}
# devtools::install_github("james-thorson/FishLife")
library(FishLife)
# remotes::install_github( 'ropensci/rfishbase@fb-21.06', force=TRUE )
library(rfishbase)
```