---
title: "Bio_parameter_calculation"
author: "Dustin Duncan"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

## Estimating catchability from abundance and cpue data

Catchability can be calculated different ways for species. One could use

\begin{align}
q_{ijm} = - [ln(-1\frac{area_{m}RET_{ijm}}{AREA})]
\end{align}

Where AREA is the area occupied by the stock, $RET_{ijm}$ is the gear retention of species *i* by age *j* by the gear of vessel type *m*, defined by a gear selection ogive, and $area_{m}$ is the area swept per day.

The specific gear retention by size is:

\begin{align}
RET_{ijm} = \frac{1}{1+e^{S1_{im}-S2_{im}L_{ij}}} 
\end{align}

\begin{align}
S1_{im} = L_{im50\%} ln \frac{3}{L_{im75\%}-L_{im50\%}} 
\end{align}

\begin{align}
S2_{im} = \frac{S1_{im}}{L_{im50\%}}
\end{align}

Where $L_{im50\%}$ and $L_{im75\%}$ are, respectively, the length at 50% and 75% retention by fleet type *m*.

If we have the slope and intercept of the maturity ogive, then we can use the length and slope to estimate $L_{im75\%}$ with the following formula:

\begin{align}
L_{im75\%} = L_{im50\%} + \frac{ln(1/3)}{a}
\end{align}

Where $L_{im50\%}$ = 13.99, and a = 12.204

```{r}
SR = 19.5 - 13.99 
L75 = 13.99 + 0.5*SR # Approximately 16.745

cpue = read.csv(here::here("data_int/data_FAO/cpue_timeseries_1990_2017.csv"))
effort = read.csv(here::here("data_int/data_FAO/effort_timeseries_1990_2017.csv"))
catch = read.csv(here::here("data_int/data_FAO/catch_timeseries_1990_2017.csv"))
ggplot(cpue, aes(x = Year, y = CPUE, color = Fleet)) + 
  geom_line()
ggplot(effort, aes(x = Year, y = Effort, color = Fleet)) + 
  geom_line()
ggplot(catch, aes(x = Year, y = Catch_tonnes, color = Fleet)) + 
  geom_line()
```

### Another method for calculating catchability 

Another method that we can use to calculate catchability is the cathc, effort, and biomass data which we have for each of the years. Maybe we can use individual years and average them, but right now we're going to just use the same catchability for either. We will use the last year of catch for industrial fisheries, and see if it works. 

Formula: 

\begin{align}
q = (C/E)/B
\end{align}

Where C is catch, E is effort, and B is biomass.

```{r}
catch_filtered = catch %>% 
  filter(Year %in% 2011)
effort_filtered = effort %>% 
  filter(Year %in% 2011)

c_e_joined = catch_filtered %>% 
  select(-c(X, Country)) %>% 
  inner_join(effort_filtered, by = c("Year", "Fleet")) %>% 
  select(-c(X, Country)) %>% 
  mutate(Biomass = 100000*0.6) %>% 
  mutate(q = ((Catch_tonnes/Effort)/Biomass))
q_ssf = c_e_joined[1, 6]
q_dwf = c_e_joined[2, 6]
options(scipen=999)
q_ssf
q_dwf

b_2010=3241/(q_ssf*139987)
effort=139/(q_dwf*b_2010)


```

Estimated that by 2017 the stock was 60,000 tonnes, so we can use this with catchability and yield from the following time-series to estimate effort for the years available in the NaFAA data? 


