---
title: "Bio_parameter_calculation"
author: "Dustin Duncan"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(devtools)
# install_github("https://github.com/haddonm/MQMF")
# devtools::install_github("https://github.com/haddonm/MQMF")
library(MQMF)
```

## Estimating catchability from abundance and cpue data

Catchability can be calculated different ways for species. One could use

\begin{align}
q_{ijm} = - [ln(-1\frac{area_{m}RET_{ijm}}{AREA})]
\end{align}

Where AREA is the area occupied by the stock, $RET_{ijm}$ is the gear retention of species *i* by age *j* by the gear of vessel type *m*, defined by a gear selection ogive, and $area_{m}$ is the area swept per day.

The specific gear retention by size is:

\begin{align}
RET_{ijm} = \frac{1}{1+e^{S1_{im}-S2_{im}L_{ij}}} 
\end{align}

\begin{align}
S1_{im} = L_{im50\%} ln \frac{3}{L_{im75\%}-L_{im50\%}} 
\end{align}

\begin{align}
S2_{im} = \frac{S1_{im}}{L_{im50\%}}
\end{align}

Where $L_{im50\%}$ and $L_{im75\%}$ are, respectively, the length at 50% and 75% retention by fleet type *m*.

If we have the slope and intercept of the maturity ogive, then we can use the length and slope to estimate $L_{im75\%}$ with the following formula:

\begin{align}
L_{im75\%} = L_{im50\%} + \frac{ln(1/3)}{a}
\end{align}

Where $L_{im50\%}$ = 13.99, and a = 12.204

```{r}
SR = 19.5 - 13.99 
L75 = 13.99 + 0.5*SR # Approximately 16.745

cpue = read.csv(here::here("data_int/data_FAO/cpue_timeseries_1990_2017.csv"))
effort = read.csv(here::here("data_int/data_FAO/effort_timeseries_1990_2017.csv"))
catch = read.csv(here::here("data_int/data_FAO/catch_timeseries_1990_2017.csv"))
ggplot(cpue, aes(x = Year, y = CPUE, color = Fleet)) + 
  geom_line()
ggplot(effort, aes(x = Year, y = Effort, color = Fleet)) + 
  geom_line()
ggplot(catch, aes(x = Year, y = Catch_tonnes, color = Fleet)) + 
  geom_line()
```

### Another method for calculating catchability 

Another method that we can use to calculate catchability is the cathc, effort, and biomass data which we have for each of the years. Maybe we can use individual years and average them, but right now we're going to just use the same catchability for either. We will use the last year of catch for industrial fisheries, and see if it works. 

Formula: 

\begin{align}
q = (C/E)/B
\end{align}

Where C is catch, E is effort, and B is biomass.

```{r}
catch_filtered = catch %>% 
  filter(Year %in% 2011)
effort_filtered = effort %>% 
  filter(Year %in% 2011)

c_e_joined = catch_filtered %>% 
  select(-c(X, Country)) %>% 
  inner_join(effort_filtered, by = c("Year", "Fleet")) %>% 
  select(-c(X, Country)) %>% 
  mutate(Biomass = 100000*0.6) %>% 
  mutate(q = ((Catch_tonnes/Effort)/Biomass))
q_ssf = c_e_joined[1, 6]
q_dwf = c_e_joined[2, 6]
options(scipen=999)
q_ssf
q_dwf

b_2010=3241/(q_ssf*139987)
effort=139/(q_dwf*b_2010)


```

Estimated that by 2017 the stock was 60,000 tonnes, so we can use this with catchability and yield from the following time-series to estimate effort for the years available in the NaFAA data? 

## Looking at whether CPUE and catch are linearly related 

```{r}
# Artisanal
cpue_s = cpue %>% 
  filter(Fleet %in% c("Artisanal")) 
catch_s = catch %>% 
  filter(Fleet %in% c("Artisanal")) 
bound_data_s = cpue_s %>% 
  full_join(catch_s, by = c("Year", "Fleet", "Country")) %>% 
  select(-c(X.x, X.y, Country)) %>% 
  filter(Year > 1997) %>% 
  mutate(Catch_tonnes = as.numeric(Catch_tonnes)) %>% 
  janitor::clean_names()

model = lm(bound_data_s$cpue ~ bound_data_s$catch_tonnes)
summary(model) # Hooray there is a correlation 

cpue_d = cpue %>% 
  filter(Fleet %in% c("Industrial")) 
catch_d = catch %>% 
  filter(Fleet %in% c("Industrial")) 
bound_data_d = cpue_d %>% 
  full_join(catch_d, by = c("Year", "Fleet", "Country")) %>% 
  select(-c(X.x, X.y, Country)) %>% 
  drop_na() %>% 
  mutate(Catch_tonnes = as.numeric(Catch_tonnes)) %>% 
  janitor::clean_names()

model2 = lm(bound_data_d$cpue ~ bound_data_d$catch_tonnes)
summary(model2) # Hooray there is a correlation (but its positive)
```

## ASPM Model from "Using R for Modelling and Quantitative Methods in Fisheries "(M. Haddon)

```{r}
# If we have natural mortality, estimated biomass, and catch, we might be able to model the population and search for the level of fishing mortality that would lead to the catch seen. If this were the case then we could then optimize the age class numbers, and estimate the ability to predict catch from fishing mortality estimates. First we will implement the age structure, and then we will implement catch to see how it works 

# Notes from Anthony, 2019: 
# Average size of sexual maturity is 20.75cm (males) and 21.76cm (females). Suggests 50% of feamles are mature at 17.50cm, and 16.60cm for males.
# Juveniles are fish that are smaller than L50, mature fish are between length L50 and Lopt, and mega-spawners are defined as fish that are Lopt plus 10% in size 
# If size can be used as a proxy for age, then it would be easy to classify the years at which certain cohorts should transfer into the above cohort, and use categorical age classes rather than actual age of the fish. 
# spm() 

# This function makes sense. I would like the output to be at FIRST, just the biological projection. After, I would like it to include the interaction between both catches, with catch being determined by effort and cost per unit effort. This will allow me to calibrate the model on the 1990-2017 data, then I should be able to back out effort from the 2018-2023 data with the information this model gives me. If I can predict effort from catch and consistent CPUE data, then I should be able to estimate effort for the new data and project it going forward under different price regimes. 
```










